---
title: "Assignment 1 ¢ñ"
author: "Xinyi Hou"
ID: "913411456"
date: "Oct,1st,2015"
output: 
  word_document: 
    fig_height: 6
    fig_width: 10
---

Load the data:

```{r}
load(url("http://eeyore.ucdavis.edu/stat141/Data/vehicles.rda"))
```

1.How many observations are there in the data set?

```{r}
nrow(vposts)
```

or

```{r}
length(vposts$id)
```

2.What are the names of the variables? and what is the class of each variable?

```{r}
names(vposts)
sapply(vposts,class)
```

3.What is the average price of all the vehicles? the median price? and the deciles? Displays these on a plot of the distribution of vehicle prices.

```{r}
avg=mean(vposts$price,na.rm=TRUE)
avg
mid=median(vposts$price,na.rm=TRUE)
mid
q<-quantile(vposts$price,seq(0,0.9,0.1),na.rm=TRUE)
q
```

As there are some extreme big values in price, I calculate the percentiles of 0.99 and omit prices bigger than it to get a more clear histogram. And I find that there is no obvious difference between the new and old median. So the abandonment of some data is reasonable.Furthermore, I plot deciles by using them as breaks of histogram.
```{r}
quantile(vposts$price,c(0.01,0.99),na.rm=TRUE)
#The 0.99 percentiles of price is 47000.
newprice<-vposts$price[vposts$price<47000]
hist(newprice,breaks=c(as.double(q),max(newprice,na.rm=TRUE)),xlim=c(0,50000),main = "Distribution of Price",xlab='Price')
points(mid,0,col='red',pch=1)
points(avg,0,col='red',pch=0)
legend(40000,1e-04,c("Median", "Mean"), pch = c(1,0),col='red')
boxplot(newprice,ylab='Price',main='Boxplot of Price')
```

4.What are the different categories of vehicles, i.e. the type variable/column? What is the proportion for each category ?

```{r}
counts=table(vposts$type)
counts
propotion=counts/sum(!is.na(vposts$type))
print(propotion,digits=2)
```

5.Display the relationship between fuel type and vehicle type. Does this depend on transmission type?

```{r}
library(ggplot2)
counts=table(vposts$fuel,vposts$type)
barplot(counts,legend=rownames(counts),col=c('red','green','yellow','blue','grey'),main='Barplot between Type and Fuel')


ggplot(vposts[!is.na(vposts$type)&!is.na(vposts$fuel)&!is.na(vposts$transmission),],aes(fuel,fill=type))+geom_bar()+facet_wrap(~transmission)+ggtitle("Relationship among Fuel,Vehicle and Transmission")
```

According to the first barplot,most of vehicles use gas as fuel. What is more, the percentages of truck,pickup and van using diesel are higher than others. This finding is the same as our common knowledge that these vehicles need more power. Hatchback use hybrid and electric as fuel more that this is a environment friendly vehicle. Last,about half of the 'other' vehicle uses 'other' fuel and these vehicles may be really special.

In the second graph,I take transmission type into considering.Automatic vehicles try all kinds of fuel, yet other transmissions are more classical which only use gas,diesel and other. When I focus on gas using vehicles, sedan and SUV take most percentages in automatic vehicles part,but there is less difference among vehicle type of manual. In other words, most sedan and SUV are automatic, yet other kinds do not show such obvious bias. 

6.How many different cities are represented in the dataset?

```{r}
unique(vposts$city)
length(unique(vposts$city))
```

7.Visually display how the number/proportion of "for sale by owner" and "for sale by dealer" varies across city?

```{r}
library(vcd)
counts<-table(vposts$byOwner,vposts$city)
counts
barplot(counts,legend=rownames(counts),col=c(87,525),main='Barplot of City and Sealer',ylab='Sold by Owner')
counts<-table(vposts$city,vposts$byOwner)
spine(counts,ylab='Sold by Owner',main='Spine Graph of Sealer and City')
```

According to the table and graphs, we can see that there is no significant difference between numbers of vehicles sold by dealer or owner among the cities. Roughly, half of the vehicles are sold by owner and the others by dealer and no matter what the city is.

8.What is the largest price for a vehicle in this data set? Examine this and fix the value. Now examine the new highest value for price.

```{r}
max(vposts$price,na.rm=TRUE)
#hist(vposts$price)
#rug(vposts$price,lwd=4)
#hist(vposts$price[vposts$price<8e4])
#rug(vposts$price[vposts$price<8e4],lwd=4)
vposts$price[vposts$price>8e4]=NA
max(vposts$price,na.rm=TRUE)
```

I use the function hist() and rug() several times to help me find extreme values of price. And I use 8e4 as the standard and fix the value by replacing all price bigger than 8e4 as NA.

9.What are the three most common makes of cars in each city for "sale by owner" and for "sale by dealer"? Are they similar or quite different?

```{r}
library(dplyr)
library(data.table)
vposts1=count(vposts,maker,city,byOwner)
vposts1=vposts1 %>% 
  filter(!maker=='NA')
#The data is ordered based on n, the number of every group we counted above.
d <- data.table(vposts1, key='n')
#.SD means subset the data, 'by' gives the method of grouping.
d<-d[, tail(.SD, 3), by=c('byOwner','city')]
arrange(d,city,byOwner)

#Another Method
#bymakercity=split(vposts1,list(vposts1$city,vposts1$byOwner))
#sapply(bymakercity,function(x)x[(x$n)>(sort(x$n,decreasing=TRUE)[4]),'maker'])
```

The three most popular makers are similar among different cities. However, there are some other interesting findings. Such as, Ford seals great in most cities except NYC and Chevrolet is quiet common except NYC and SFbay. What is more, Honda is sold by owners more than dealers. Dodge is welcomed in Denver and BMW is common in SFbay,yet they are not three top popular makers in other cities.

Reference:
http://stackoverflow.com/questions/14800161/how-to-find-the-top-n-values-by-group-or-within-category-groupwise-in-an-r-dat

10.Visually compare the distribution of the age of cars for different cities and for "sale by owner" and "sale by dealer". Provide an interpretation of the plots, i.e., what are the key conclusions and insights?

```{r}
library(dplyr)
library(ggplot2)
vposts2=
  vposts %>% 
  filter(year<=2015&year>1800) %>% 
  mutate(age=2015-year)
ggplot(vposts2,aes(age, fill = byOwner)) + geom_histogram(alpha = 0.5)+ggtitle("Histogram  of Age of Different Sealer") 
ggplot(vposts2,aes(x=city,y=age,fill=city))+geom_boxplot()+ggtitle("Boxplot of Age of Different City")
vposts2 %>% 
  filter(age<60) %>% 
  ggplot(aes(age))+geom_histogram(fill='red')+facet_wrap(~city)+ggtitle('Histograms of Age of Different City')
```

Based on the histogram of age of different sealers, the age of vehicles sold by owner is longer than its sold by dealer. What we can get from the second graph is that the age medians of Denver,SAC and SFbay are bigger than others. And the ranges of age of Denver and SFbay are wider. The finding about age range can also be gotten from the histograms of age of different city.  

Reference:
http://stackoverflow.com/questions/3541713/how-to-plot-two-histograms-together-in-r

11.Plot the locations of the posts on a map? What do you notice?

```{r}
library(maps)
map(database='state')
with(vposts,points(long,lat,pch='.',col='red'))
title(main="Where does vehicle be sold?")
```

I have two findings from the map.Firstly, fewest vehicles are sold in Las Vegas compared to other cities. What is more, the vehicles are not sold only in the listed 7 cities and there are vehicles sold in other places, although just a few.

12.Summarize the distribution of fuel type, drive, transmission, and vehicle type. Find a good way to display this information.

```{r}
library(ggplot2)
library(dplyr)
vposts %>% 
  filter(!is.na(fuel))%>%
  count(fuel) %>% 
  ggplot(aes(x=1,y=n,fill=fuel))+geom_bar(stat='identity',width=1)+coord_polar("y",start=0)   +scale_fill_brewer(palette="Blues")+ggtitle('Pie Chart of Fuel Type')

#Another kind of code
#vposts %>% 
#  filter(!is.na(fuel))%>%
#  ggplot(aes(x=factor(1),fill=fuel))+geom_bar(width=1)+coord_polar("y",start=0)   #+scale_fill_brewer(palette="Blues")

vposts %>% 
  filter(!is.na(drive))%>%
  ggplot(aes(drive))+geom_bar(aes(fill=drive))+ggtitle('Bar Chart of Drive') 

vposts %>% 
  filter(!is.na(transmission))%>%
  ggplot(aes(x=factor(1),fill=transmission))+geom_bar()+ggtitle('Bar Chart of Transmission')

vposts %>% 
  filter(!is.na(type))%>%
  ggplot(aes(type))+geom_bar(aes(fill=type))+ggtitle('Bar Chart of Vehicle Type')

vposts %>% 
  filter(!is.na(type)&!is.na(fuel)&!is.na(transmission)&!is.na(drive))%>%
  ggplot(aes(fuel,fill=type))+geom_bar()+facet_wrap(drive~transmission)+ggtitle("Relationship among Fuel,Vehicle,Drive and Transmission")
```

As we know, most of vehicles are automatic, front-wheel drive and use gas as fuel. And I notice that most suv is 4wd, yet most sedan is fwd.  

Reference:
http://www.sthda.com/english/wiki/ggplot2-pie-chart-quick-start-guide-r-software-and-data-visualization

13.Plot odometer reading and age of car? Is there a relationship? Similarly, plot odometer reading and price? Interpret the result(s). Are odometer reading and age of car related?

```{r}
library(ggplot2)
library(dplyr)
#To get a good graph, I omit some obeservations whose odometers are extreme big and age are old.
vposts2 %>% 
  filter(!is.na(odometer)&(odometer<5e5)&(age<30)) %>% 
  ggplot(aes(x=odometer,y=age))+geom_point(alpha=0.3)+geom_smooth(size=1.5)+ggtitle('Point Plot of Odometer Reading and Age ')

vposts2 %>% 
  filter(!is.na(odometer)&(odometer<3e5)&!is.na(price)&(price<8e4)) %>% 
  ggplot(aes(x=price,y=odometer))+geom_point(alpha=0.3)+geom_smooth(size=1.5)+ggtitle('Point Plot of Odometer Reading and Price ')
```

According to the plot between age and odometer reading, there is a positive relationship between them in the concentrated area. Instead, there is a negative relationship between price and odometer reading in the concentrated area. Actually, in my opoinion, these relationships are not very obvious.

14.Identify the "old" cars. What manufacturers made these? What is the price distribution for these?

Atomakers say they are building the best cars ever, and there¡¯s proof on the road. The average age of vehicles in the U.S. has climbed to a record 11.5 years, according to research firm IHS Automotive. So I define the old car as the car whose age is bigger than or equals to 12 years. 

```{r}
library(dplyr)
vposts2 %>% 
  filter(age>=12) %>% 
  count(maker) %>% 
  arrange(desc(n))

#Use the same threshhold of price as queation 8. 
vposts2 %>% 
  filter(age>=12&price<8e4) %>% 
  ggplot(aes(price,alpha=0.5))+geom_histogram(aes(fill='red'))+ggtitle('Distribution of Old Car')
```

Reference:
http://www.latimes.com/business/autos/la-fi-hy-ihs-average-car-age-20150729-story.html

15.I have omitted one important variable in this data set. What do you think it is? Can we derive this from the other variables? If so, sketch possible ideas as to how we would compute this variable.

As far as I see, MPG(miles per gallon) is missed. However,professor Duncan said on piazza that only about 1500 observations contain the MPG in the 'body' and it is not enough to construct another variable. In addition, the color is also missed. Someone posted on piazza that a lot of observations contain this information in the 'body' and it is possible to extract out and build another variable. 

16.Display how condition and odometer are related. Also how condition and price are related. And condition and age of the car. Provide a brief interpretation of what you find.
If you display results in a table, consider a more appropriate plot.
If you do use a table, consider the number of digits in the numbers displayed.
If you use a single plot, consider using a legend to identify the different sub-elements.
If you use multiple panels/plots, consider if the scales are the same and ensure they are if appropriate.
Make certain plots have the approriate labels on the axes.

```{r}
library(ggplot2)
library(dplyr)

#I only plot the conditions which show more than 5 times in the dataframe as the drop part is not representative enough. After, we can get a more clear graph.
commoncondition=
  vposts %>% 
  filter(!is.na(condition)) %>% 
  count(condition) %>% 
  filter(n>5)

#To compare easily,omit some large odometer cars 
vposts %>% 
  filter(!is.na(odometer)&(odometer<1e6)&!is.na(condition)& (condition %in%     commoncondition$condition)) %>% 
  ggplot(aes(x=condition,y=odometer))+geom_boxplot()+ggtitle("Boxplot of Condition and Odometer Reading")

vposts %>% 
  filter(!is.na(price)&!is.na(condition)& (condition %in% commoncondition$condition)) %>% 
  ggplot(aes(x=condition,y=price))+geom_boxplot()+ggtitle("Boxplot of Condition and Price")

vposts2 %>% 
  filter(!is.na(condition)& (condition %in% commoncondition$condition)) %>% 
  ggplot(aes(x=condition,y=age))+geom_boxplot()+ggtitle("Boxplot of Condition and Age")

```

We can get some interesting results from three graphs. Mostly,the vehicles in positive conditions such as 'like new' or 'excellent' have higher price,lower odometer reading and shorter age comapred to vehicles in the general or negative conditions such as 'fair' or 'salvage'. In addition, certified and used vehicles have highest price, lowest age and almost lowest odometer readings. However, there are a lot of extreme big values of odometer readings on excellent,good and like new condition. There must be some sealers using better words to describe cars and it is a little faked. Besides, the condition 'very good' and 'fair' have similar odometer reading and age, yet the price of 'very good' vehicles are higher. Thus, some vehicles in 'very good' condition may not work so well actually.   
